---
title: "R Notebook"
output: html_notebook
---
Install Library package and data
```{r}
library(ggplot2)
require(ggrepel)
require(dplyr)
require(stats)
country_waste_data <- read.csv("country_level_data.csv")
city_waste_data <- read.csv("city_level_data.csv")
municipal_waste_data <- read.csv("municipal_waste.csv")
```

Renaming columns
```{r}
colnames(country_waste_data)[6] <- "organic_waste"
colnames(country_waste_data)[7] <- "glass_waste"
colnames(country_waste_data)[8] <- "metal_waste"
colnames(country_waste_data)[9] <- "other_waste"
colnames(country_waste_data)[10] <- "paper_cardboard_waste"
colnames(country_waste_data)[11] <- "plastic_waste"
colnames(country_waste_data)[12] <- "rubber_&_leather_waste"
colnames(country_waste_data)[13] <- "wood_waste"
colnames(country_waste_data)[14] <- "green_waste"
country_waste_data
```

```{r}
        ggplot(country_waste_data, aes(x = country_name, y = 'plastic waste')) +
          geom_col() +
          geom_bar(stat="identity", fill = "darkred") +
          xlab("Country") + ylab("total waste in tonnes")
```

Dataset inspection.
Think of questions to be asked.
Check if column contains null values.
```{r}
dim(country_waste_data)
str(country_waste_data)
summary(country_waste_data)
colnames(country_waste_data)
is.na(country_waste_data$total_msw_total_msw_generated_tons_year)
```

Top 10 highest waste generating countries
```{r}
complete <- na.omit(country_waste_data$total_msw_total_msw_generated_tons_year)
sorted <- country_waste_data[order(-country_waste_data$total_msw_total_msw_generated_tons_year),]
top_10 <- head(sorted, 10)
ggplot(top_10, aes(x = reorder(country_name, +total_msw_total_msw_generated_tons_year), y = total_msw_total_msw_generated_tons_year)) +
  geom_col() + 
  geom_bar(stat="identity", fill = "darkred") +
  xlab("Country") + ylab("total waste in tonnes") + 
  coord_flip()
```

Global waste composition
```{r}
sorted <- country_waste_data[order(country_waste_data$organic_waste, decreasing=TRUE),]
sorted
top_10 <- head(sorted, 10)

ggplot(top_10, aes(x = reorder(country_name, +organic_waste), y = organic_waste)) +
  geom_col() + 
  geom_bar(stat="identity", fill = "darkred") +
  xlab("Country") + ylab("total waste in tonnes") + 
  coord_flip()
```

What constitutes each country's waste? 
Pie Chart. Can select which country's pie chart you want.
```{r}

```

Calculate GDP per capita
```{r}
gpc <- country_waste_data$gdp/country_waste_data$population_population_number_of_people
country_waste_data$gdp_per_capita <- gpc
country_waste_data$gdp_per_capita
```

waste generation calculate per kg
```{r}
country_waste_data$total_waste_per_kg <- country_waste_data$total_msw_total_msw_generated_tons_year*1000
country_waste_data$total_waste_per_kg
```

Calculate waste generation per capita
Waste Generated divided by population
```{r}
country_waste_data$waste_per_capita <- country_waste_data$total_waste_per_kg/country_waste_data$population_population_number_of_people
country_waste_data$waste_per_capita_day <- country_waste_data$waste_per_capita/365
country_waste_data$waste_per_capita_day
```

Scatter plot with population

```{r}
cdat <- country_waste_data %>% mutate(plotname = as.character(country_name))
cdat <- country_waste_data %>% mutate(plotname = as.character(country_name))

countrylist <- c("Sierra Leone", "Argentina", "Canada", "Ireland", "United Kingdom", "United States",
  "New Zealand", "Iceland", "Japan", "Luxembourg", "Netherlands", "Switzerland", "Qatar", "South Africa", "Rwanda", "China")

cdat <- cdat %>%
  mutate(plotname = ifelse(plotname %in% countrylist, plotname, ""))

cdat %>%
  select(country_name, plotname)
```

```{r}
ggplot(cdat, aes(x = population_population_number_of_people, y = waste_per_capita)) +
  geom_point() +
  xlab("population") + 
  ylab("waste per capita (kg/person/year)") +
  geom_text_repel(aes(label = plotname), size = 4)
```

Calculate average total waste for each income group
```{r}
hic <- country_waste_data %>% filter(income_id == "HIC")
umc <- country_waste_data %>% filter(income_id == "UMC")
lmc <- country_waste_data %>% filter(income_id == "LMC")
lic <- country_waste_data %>% filter(income_id == "LIC")

hic_mean <- mean(hic$total_msw_total_msw_generated_tons_year, na.rm = T)
umc_mean <- mean(umc$total_msw_total_msw_generated_tons_year, na.rm = T)
lmc_mean <- mean(lmc$total_msw_total_msw_generated_tons_year, na.rm = T)
lic_mean <- mean(lic$total_msw_total_msw_generated_tons_year, na.rm = T)

hic_pop <- mean(hic$population_population_number_of_people, na.rm = T)
umc_pop <- mean(umc$population_population_number_of_people, na.rm = T)
lmc_pop <- mean(lmc$population_population_number_of_people, na.rm = T)
lic_pop <- mean(lic$population_population_number_of_people, na.rm = T)

hic_mean_capita <- mean(hic$waste_per_capita_day, na.rm = T)
umc_mean_capita <- mean(umc$waste_per_capita_day, na.rm = T)
lmc_mean_capita <- mean(lmc$waste_per_capita_day, na.rm = T)
lic_mean_capita <- mean(lic$waste_per_capita_day, na.rm = T)

hic_gdp <- mean(hic$gdp, na.rm = T)
umc_gdp <- mean(umc$gdp, na.rm = T)
lmc_gdp <- mean(lmc$gdp, na.rm = T)
lic_gdp <- mean(lic$gdp, na.rm = T)



id_mean <- c(hic_mean, umc_mean, lmc_mean, lic_mean)
gdp_mean <- c(hic_gdp, umc_gdp, lmc_gdp, lic_gdp)
capita_mean <- c(hic_mean_capita, umc_mean_capita, lmc_mean_capita, lic_mean_capita)

id_mean_df <- data.frame(id_mean, gdp_mean, capita_mean)
id_mean_df
```

Bubble plot
```{r}
ggplot(id_mean_df, aes(x=gdp_mean, y=capita_mean, size = id_mean)) +
    geom_point(alpha=0.8) +
    scale_size(range = c(1, 15), name="Total waste in millions")
```
Donut dataframe
```{r}
a <- c(sum(country_waste_data$total_msw_total_msw_generated_tons_year, na.rm = T))
b <- c("total waste in millions per tonne")
donut <- data.frame(a, b)
donut$labelPosition <- (donut$a + 0) / 2
donut$label <- paste0(donut$b, "\n value: ", donut$a)

ggplot(donut, aes(ymax=a, ymin=0, xmax=4, xmin=3, fill=b), scientific = FALSE) +
     geom_rect() +
     geom_label( x=3.5, aes(y=labelPosition, label=label), size=3) +
     coord_polar(theta="y") + 
     xlim(c(2, 4)) +
     theme_void() +
     theme(legend.position = "none")
```

Donut Plot
```{r}
LCN <- country_waste_data %>% filter(region_id == "LCN")
SAS <- country_waste_data %>% filter(region_id == "SAS")
SSF <- country_waste_data %>% filter(region_id == "SSF")
ECS <- country_waste_data %>% filter(region_id == "ECS")
EAS <- country_waste_data %>% filter(region_id == "EAS")
MEA <- country_waste_data %>% filter(region_id == "MEA")
NAC <- country_waste_data %>% filter(region_id == "NAC")

total_LCN <- sum(LCN$total_msw_total_msw_generated_tons_year, na.rm = T)
total_SAS <- sum(SAS$total_msw_total_msw_generated_tons_year, na.rm = T)
total_SSF <- sum(SSF$total_msw_total_msw_generated_tons_year, na.rm = T)
total_ECS <- sum(ECS$total_msw_total_msw_generated_tons_year, na.rm = T)
total_EAS <- sum(EAS$total_msw_total_msw_generated_tons_year, na.rm = T)
total_MEA <- sum(MEA$total_msw_total_msw_generated_tons_year, na.rm = T)
total_NAC <- sum(NAC$total_msw_total_msw_generated_tons_year, na.rm = T)

region <- c("LCN", "SAS", "SSF", "ECS", "EAS", "MEA", "NAC")
waste <- c(total_LCN, total_SAS, total_SSF, total_ECS, total_EAS, total_MEA, total_NAC)

total_waste <- data.frame(region, waste)

total_waste$fraction <- total_waste$waste / sum(total_waste$waste)

total_waste$ymax <- cumsum(total_waste$fraction)
total_waste$ymin <- c(0, head(total_waste$ymax, n=-1))

total_waste$labelPosition <- (total_waste$waste + 0) / 2
total_waste$label <- paste0(total_waste$region, "\n value: ", total_waste$region)

ggplot(total_waste, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=region), scientific = FALSE) +
     geom_rect() +
     # geom_text( x=3.0, aes(y=labelPosition, label=label), size=3) +
     coord_polar(theta="y") + 
     xlim(c(2, 4))  
     # theme_void()
     # theme(legend.position = "none")
```

```{r}
hic <- country_waste_data %>% filter(income_id == "HIC")
umc <- country_waste_data %>% filter(income_id == "UMC")
lmc <- country_waste_data %>% filter(income_id == "LMC")
lic <- country_waste_data %>% filter(income_id == "LIC")

total_HIC <- sum(hic$total_msw_total_msw_generated_tons_year, na.rm = T)
total_UMC <- sum(umc$total_msw_total_msw_generated_tons_year, na.rm = T)
total_LMC <- sum(lmc$total_msw_total_msw_generated_tons_year, na.rm = T)
total_LIC <- sum(lic$total_msw_total_msw_generated_tons_year, na.rm = T)

income <- c("HIC", "UMC", "LMC", "LIC")
income_total_waste <- c(total_HIC, total_UMC, total_LMC, total_LIC)

income_waste <- data.frame(income, income_total_waste)

income_waste
income_waste$fraction <- income_waste$income_total_waste / sum(income_waste$income_total_waste)

income_waste$ymax <- cumsum(income_waste$fraction)
income_waste$ymin <- c(0, head(income_waste$ymax, n=-1))

income_waste$labelPosition <- (income_waste$income_total_waste + 0) / 2
income_waste$label <- paste0(income_waste$income, "\n value: ", income_waste$income)

ggplot(income_waste, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=income), scientific = FALSE) +
     geom_rect() +
     # geom_text( x=3.0, aes(y=labelPosition, label=label), size=3) +
     coord_polar(theta="y") + 
     xlim(c(2, 4))  
     # theme_void()
     # theme(legend.position = "none")
```

Animated Pie Chart
```{r}
# library(gganimate)
# library(gifski)
# library(png)
# 
# cv <- c(sum(country_waste_data$total_msw_total_msw_generated_tons_year, na.rm = T))
# # Make 2 basic states and concatenate them:
# a <- data.frame(waste=c("World"), total_waste_in_the_world_per_million_tonnes = 0, frame=rep('a',3))
# b <- data.frame(waste=c("World"), total_waste_in_the_world_per_million_tonnes =cv, frame=rep('b',3))
# data <- rbind(a,b)  
#  
# # Basic barplot:
# ggplot(a, aes(x=waste, y=total_waste_in_the_world_per_million_tonnes, fill=waste), scientific = FALSE) + 
#   geom_bar(stat='identity')
#  
# # Make a ggplot, but add frame=year: one image per year
# gg <- ggplot(data, aes(x=waste, y=total_waste_in_the_world_per_million_tonnes, fill=waste), scientific = FALSE) + 
#   geom_bar(stat='identity', scientific = FALSE) +
#   theme_bw() +
#   # gganimate specific bits:
#   transition_states(
#     frame,
#     transition_length = 10,
#     state_length = 1
#   ) +
#   ease_aes('sine-in-out')
# 
# # Save at gif:
# animate(gg, end_pause = 10, width=1000, height=600,fps = 5)
# anim_save("288-animated-barplot-transition.gif", renderer = gifski_renderer())
```

Total Plot
```{r}
wt <- sum(country_waste_data$total_msw_total_msw_generated_tons_year, na.rm = T)
ht <- sum(country_waste_data$special_waste_hazardous_waste_tons_year, na.rm = T)
it <- sum(country_waste_data$special_waste_industrial_waste_tons_year, na.rm = T)

total_alarm <- data.frame(waste = c("total_waste_in_the_world", "total_hazardous_waste", "total_industrial_waste"), waste_in_million_tonnes = c(wt, ht, it))

# total_alarm <- data.frame(total_waste_in_the_world = c(wt), total_hazardous_waste = c(ht), total_industrial_waste = c(it))

total_alarm

ggplot(total_alarm, aes(x = waste, y = total_waste_in_the_world)) +
  geom_col() + 
  geom_bar(stat="identity", fill = "darkred") +
  xlab("Alarming Wastes") + ylab("total waste in million tonnes")
```
```{r}
compareSelect <- country
```

```{r}
compareSelect <- country_waste_data %>% select(glass_waste, organic_waste)
compareSelect
colSums(compareSelect, na.rm = T)
comparedf <- data.frame(c)
sumCompare <- sum(c(compareSelect), na.rm = T)
sumCompare

# Calculate the sum of all waste types first.
# Put it into a new df.
# Then if triggered, select the columns accordingly.
```

municiipal waste treatment
```{r}
treated <- municipal_waste_data %>% filter(Variable == "Municipal waste treated", Country == "Australia")
treated2 <- treated %>% select(Year, Value)
ggplot(treated2, aes(x=Year, y=Value), scientific = FALSE) +
  geom_line() +
  geom_point()
```








